---
title: "Project-Stage 2"
author: "Katy Schmitt"
date: "2023-10-30"
output: html_document
---

```{r, echo=FALSE}
install.packages(c("tidyverse", "ggplot2", "readxl", "binsreg", "tidyr", "dplyr", "sandwich", "RColorBrewer"))
install.packages("binscatter")
library(tidyverse)
library(ggplot2)
library(RColorBrewer)
library(readxl)
library(binsreg)
library(tidyr)
library(dplyr)
library(sandwich)
Master_data_2 <- read_excel("~/Master_data-2.xlsx")
#View(Master_data_2)

```

## Immigrant Populations and Presidential Election Outcomes ##
### Katy Schmitt and Md Sayemul Islam ###
### Stage 3&4 Graph Creation ###

### Advanced Data Analytics - ECNS_560 ###
#### December 2, 2023 ####  

#### **Female** Findings ####

1) How does **female associate with Vote Share Difference**?
```{r, echo=FALSE}
Scatter1 = ggplot(data=Master_data_2, aes(Master_data_2$Female, Master_data_2$Vote_share_difference)) +geom_point(color="pink") +labs(title ='Female and Vote Share Difference', x='Female Population (%)', y='Vote Share Difference') + theme_classic() +geom_hline(yintercept=0, linetype="dashed", color = "black") +geom_smooth()
Scatter1
```


2) Distribution of Foreign Born population in US counties
```{r}
#adding a pdf image of pretty hist here

Foreign_Hist = ggplot(data=Master_data_2, aes(x=Foreign_born))+
  geom_histogram(bins=40, color="light blue")+
  labs(x= 'Foreign Born Pop. (%)', y='County Frequency', title = 'Foreign Born Distribution in US Counties') +theme_classic()

Foreign_Hist
```



3) California, New York, and Texas Vote Share Trends Over Time

```{r, echo=FALSE}
years = c(2012, 2016, 2020)
states = c("CALIFORNIA", "TEXAS","NEW YORK")
Master_data_2  %>%
  filter(Year %in% years & State %in% states) %>%
  ggplot( aes(Foreign_born,Vote_share_difference, col = State)) +
    geom_point() +
    facet_wrap(. ~ Year)+labs(title ='Foreign Born Population and Vote Share by State', x='Foreign Born Population (%)', y='Vote Share Difference') +geom_hline(yintercept=0, linetype="dashed", color = "black")  + theme_classic()
```


4)  Is there a relationship between **urban counties, rural counties, and foreign** born population?


```{r, echo=FALSE}
Master_data_2020 <- Master_data_2 %>%
  filter(Year==2020) 

Master_data_20202 = Master_data_2020 %>%
  group_by(State, Rural) %>%
  summarise(count = n(), Foreign_Avg = sum(Foreign_born)) #%>%

Master_data_20202 <- Master_data_20202 %>%
  mutate(state_avg = Foreign_Avg/count)
  
  
  Avg_Graph = ggplot(aes(x=State, y = state_avg), data=Master_data_20202) + geom_bar(stat = 'identity', fill= 'light blue',) + facet_wrap(~Rural) + theme_classic() +labs(title='Avg. Foreign Born Population in Urban (0) vs. Rural (1) Counties in 2020', y='Foreign Born Population Average in State (%)', x='State') +theme(text = element_text(size = 7.5)) +coord_flip()

  Avg_Graph
```

5) Examine any relationship between **foreign born population and Vote Share Difference**.
```{r, echo=FALSE}
FB_VS = ggplot(data=Master_data_2, aes(x = Foreign_born, y=Vote_share_difference)) +geom_point(show.legend = FALSE, alpha = 0.2, color = "light blue")  +labs(x = 'Foreign born Population (%)', y='Vote Share Difference', title = 'Foreign Born Population and Vote Share Difference') +geom_smooth(color="red") +geom_hline(yintercept=0, linetype="dashed", color = "black")  + theme_classic()

FB_VS
```

